---
title: "Assignment 5 - Decision Trees"
author: "Abbey Keener"
date: "November 16, 2017"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}
install.packages("party", "rpart")

library(rpart)
library(party)
```

## Part I
```{r}
D1 <- read.csv("/Users/abbeykeener/HUDK Assignments/Assignment 5/intelligent_tutor.csv", sep=",", header=TRUE)
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula X ~ Y 

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

```
## Part II

#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Take a look at our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action"))
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
```

#Plot tree
```{r}
plot(score_ctree)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?
1. Whether students asked for hints or not (and how many hints they asked for). This is the most significant relationship at p<0.001. 
2. How many problems they have completed before. Nodes 3 and 4 have big differences between intervene and no action and node 3 has the biggest N, and node 4 which shows people who didn't ask for a hint and completed more than 85 prior problems has the most students which need less action so the teacher should may closer attention to other students. 

#Test Tree
Upload the data "intelligent_tutor_new.csv". This is a data set of a differnt sample of students doing the same problems in the same system. We can use the tree we built for the previous data set to try to predict the "advice" we should give the teacher about these new students. 

```{r}
#Upload new data

D2 <- read.csv("/Users/abbeykeener/HUDK Assignments/Assignment 5/intelligent_tutor_new.csv", sep=",", header=TRUE)

#Generate predicted advice for new students based on tree generated from old students

D2$prediction <- predict(score_ctree, D2)

``` 
## Part III
Compare the predicted advice with the actual advice that these students recieved. What is the difference between the observed and predicted results?

```{r}
D2$advice <- ifelse(D2$score <=0.4, "intervene", ifelse(D2$score > 0.4 & D2$score <=0.8, "monitor", "no action"))
#creating a new advice column in D2 independent of prediction using teachers rationale from D1
#no action for all of them because they all have scores of 1 

confMat <- table(D2$prediction,D2$advice)
#shows frequency difference between predicted and actual advice using a confusion matrix 
#predictions based on D1 model compared to advice based on D2 scores and score perameters from D1

View(confMat)
#because all advice is now no action because the scores are all 1, the predicted advice was monitor and actual was no action 84 times and it was accurate 116 times (where predicted and actual were both no action) it never predicted intervene 
#Var1=predicted Var2=advice 

```
